---
title: "compsciX 415.2 Homework 3"

output: 
html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mdsr)
library(tidyverse)

"'{r load_packages, warning=FALSE, message=FALSE}"

library(ggplot2)
library(nycflights13)
data("flights")
```
```{r}
glimpse (flights)
```


```{r, fig.align='center', fig.asp=.1}

```
```{r}
ggplot(data = mpg, mapping=aes(x=cty, y=hwy)) + 
  geom_point()

ggplot(data = mpg, mapping=aes(x=cty, y=hwy)) + 
  geom_point(position = "jitter")


```

###Sect. 3.8.1  1) geom_point() plots points that can overlap so you cannot see all of data points. Using geom_point() with position="jitter" will plot dots off one another by adding a small amount of noise for each dot.

```{r}
ggplot(data = mpg, mapping=aes(x=cty, y=hwy)) + 
  geom_jitter(width=0.6, height=0.6)

ggplot(data = mpg, mapping=aes(x=cty, y=hwy)) + 
  geom_jitter(width=0.2, height=0.2)

```
###Sect 3.8.1 2) In geom_jitter(), width= and height= control the amount of vertical and horizontal jitter.





```{r}
ggplot(data = mpg, mapping=aes(x=cty, y=hwy)) + 
  geom_jitter()

ggplot(data = mpg, mapping=aes(x=cty, y=hwy)) + 
  geom_count()

```
###Sect 3.8.1 3) Geom_jitter() plots the dots slighty off one another. Geom_count() plots the dots in the same location like geom_point() so they overlap but the dots get larger if there are more dots in the same location. 
  

```{r}
ggplot(data = mpg) + 
  geom_boxplot(mapping = aes(x = cyl , y = hwy, group=class))
```
###Sect 3.8.1 4)

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = manufacturer, y = cty)) +
  labs(x="Car manufacturers", y="City mileage", title="Car manufacturers by city mpg")
```

###Sect 3.9.1 2) Labs() modifies axis, legend and plot labels, including title, caption. 


```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy))+
  geom_point() +
  geom_abline() +
  coord_fixed()
```

###Section 3.9.1  4) This plot shows that the better the city mileage, the better the highway mileage. They have a positive relationship, which is expected. Coord_fixed() sets the coordinate system of the data on the axes. The default, ratio = 1, ensures that one unit on the x-axis is the same length as one unit on the y-axis. Without this, the distance between the X, Y units can be different, skewing the visual display of the data. Geom_abline() draws the line along the data with the slope and intercept. 

```{r}
my_variable <- 10 


```

```{r}
my_variable <- 10 
my_variable

```
###Sect. 4.4 1) The original code did not work because the letter "i" in "my_variable" was incorrectly typed as "1".



```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) 
filter (mpg, cyl==8)
```

###Sect. 4.4 2) Draw a scatterplot with X is displ and Y is highway mileage for all vehicles that are 8 cylinders.


```{r}
glimpse(flights)
  
```
###Sect. 5.2.4  Code for exercises 1) 1-7

```{r}
filter(flights, arr_delay>=120)

filter(flights, dest=="IAH" | dest=="HOU")

filter(flights, carrier=="UA" | carrier=="AA" | carrier=="DL")

filter(flights, month==7 | month==8 | month==9)

filter(flights, arr_delay>=120, dep_delay<=0)

filter(flights, dep_delay>=60, arr_delay<=30)

filter(flights, dep_time>=000, dep_time<=600)
```
###Sect. 5.2.4  2) In the dply filtering option, between() allows you to filter data inclusively within a range between n and m.


```{r}
flights %>% filter(between(month, 6, 9))

flights %>% filter(between(dep_time, 000, 600))

```

```{r}
filter(flights, is.na(dep_time))

```
###Sect.5.2.4 3) 8255 flights have missing departure times. Missing are also dep_delay, arr_time, arr_delay. These are flights that did not take flight. 

###Sect. 5.2.4 4) NA ^ 0 is not missing because you have set NA to 0, which has a value of 0. NA | TRUE is not missing because NA or TRUE includes TRUE, therefore not missing. FALSE and NA is not missing because it includes FALSE. Missing is only missing if it does not intersect with anything else.



```{r}
 select(flights, dep_time, dep_delay, arr_time, arr_delay)

select(flights, dep_time:arr_delay)
```
###Sect. 5.4.1 1) You can select the variables you want directly and list them, or select them within a range of variables, or as partial names like dep_ or arr_.

```{r}
select(flights, origin, carrier, flight, origin, dest, origin)
```
###Sect. 5.4.1. 2) If you include the variable multiple times in select(), it will output the first time the variable appears but not in subsequent times.


```{r}
vars<-c("year", "month", "day", "dep_delay", "arr_delay")
select(flights,one_of(vars))

```

###Sect. 5.4.1. 3) one_of() is part of the helper function of select(). It selects any variables that you have already defined as "vars" in the above example. It is good to use once you have created a batch of variables and selected/renamed them. 


```{r}
select(flights, contains("TIME"))
```

###Sect. 5.4.1 4) select(contains()) allows you to select any variable with "time" in its name, whether in the beginning or end of the variable name.




