---
title: "compsciX 415.2 Homework 3"

output: 
html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mdsr)
library(tidyverse)

"'{r load_packages, warning=FALSE, message=FALSE}"

library(ggplot2)
library(nycflights13)
data("flights")
```
```{r}
glimpse (flights)
```


```{r, fig.align='center', fig.asp=.1}

```
```{r}
ggplot(data = mpg, mapping=aes(x=cty, y=hwy)) + 
  geom_point()

ggplot(data = mpg, mapping=aes(x=cty, y=hwy)) + 
  geom_point(position = "jitter")


```

### Sect. 3.8.1 
### 1) geom_point() plots points that can overlap so you cannot see all of data points. Using geom_point() with position="jitter" will plot dots off one another by adding a small amount of noise for each dot.

```{r}
ggplot(data = mpg, mapping=aes(x=cty, y=hwy)) + 
  geom_jitter(width=0.6, height=0.6)

ggplot(data = mpg, mapping=aes(x=cty, y=hwy)) + 
  geom_jitter(width=0.2, height=0.2)

```
###Sect 3.8.1 
###2) In geom_jitter(), width= and height= control the amount of vertical and horizontal jitter.





```{r}
ggplot(data = mpg, mapping=aes(x=cty, y=hwy)) + 
  geom_jitter()

ggplot(data = mpg, mapping=aes(x=cty, y=hwy)) + 
  geom_count()

```
###Sect 3.8.1 
###3) Geom_jitter() plots the dots slighty off one another. Geom_count() plots the dots in the same location like geom_point() so they overlap but the dots get larger if there are more dots in the same location. 
  

```{r}
ggplot(data = mpg) + 
  geom_boxplot(mapping = aes(x = cyl , y = hwy, group=class))
```
###Sect 3.8.1 
###4)

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = manufacturer, y = cty)) +
  labs(x="Car manufacturers", y="City mileage", title="Car manufacturers by city mpg")
```

###Sect 3.9.1 2) Labs() modifies axis, legend and plot labels, including title, caption. 


```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy))+
  geom_point() +
  geom_abline() +
  coord_fixed()
```

###Section 3.9.1  4) This plot shows that the better the city mileage, the better the highway mileage. They have a positive relationship, which is expected. Coord_fixed() sets the coordinate system of the data on the axes. The default, ratio = 1, ensures that one unit on the x-axis is the same length as one unit on the y-axis. Without this, the distance between the X, Y units can be different, skewing the visual display of the data. Geom_abline() draws the line along the data with the slope and intercept. 

```{r}
my_variable <- 10 


```

```{r}
my_variable <- 10 
my_variable

```
###Sect. 4.4 1) The original code did not work because the letter "i" in "my_variable" was incorrectly typed as "1".

```{r}
library(tidyverse)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) 
filter (mpg, cyl==8)
```
###Sect. 4.4 2) Draw a scatterplot with X is displ and Y is highway mileage for all vehicles that are 8 cylinders.


```{r}
library(ggplot2)
ggplot(data = flights)
glimpse(flights)
  
```
###Sect. 5.2.4  Code for exercises 1) 1-7

```{r}
filter(flights, arr_delay>=120)

filter(flights, dest=="IAH" | dest=="HOU")

filter(flights, carrier=="UA" | carrier=="AA" | carrier=="DL")

filter(flights, month==6 | month==7 | month==8)

filter(flights, arr_delay>=120, dep_delay<=0)

filter(flights, dep_delay>=60)

filter(flights, dep_time>=000, dep_time<=600)
```
###Sect. 5.2.4  2) In the dply filtering option, between() allows you to filter data inclusively within a range between n and m.


```{r}
flights %>% filter(between(month, 6, 9))

flights %>% filter(between(dep_time, 000, 600))

```



```{r}
View(flights)
```
```{r}
filter(flights, is.na(dep_time))

```
###Sect.5.2.4 3) 8255 flights have missing departure times. Missing are also dep_delay, arr_time, arr_delay. These are flights that did not take flight. 

###Sect. 5.2.4 4)



```{r}
 select(flights, dep_time, dep_delay, arr_time, arr_delay)

select(flights, dep_time:arr_delay)
```

```{r}
select(flights, origin, carrier, flight, origin, dest, origin)
```
###Sect. 5.4.1. 2) If you include the variable multiple times in select(), it will output the first time the variable appears but not in subsequent times.


```{r}
(vars<-c("year", "month", "day", "dep_delay", "arr_delay"))

vars %>% select(one_of(c("year", "month", "day", "dep_delay", "arr_delay")))
```


vars
```

mtcars %>% select(one_of(c("mpg","cyl","garbage")))


