---
title: "Homework_2_Quynh_Tran"
output:
  html_document: default
  
---

"'{r load_packages, warning=FALSE, message=FALSE}"

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mdsr)
library(tidyverse)
library(ggplot2)

data("mpg")
```

```{r, fig.align='center', fig.asp=.1}
ggplot(data=mpg)

```
###Section 3.2.4 1) You see nothing with ggplot(data=mpg)
###Section 3.2.4 2) There are 234 rows and 11 columns
###Section 3.2.4 3) Variable drv describes drivetrain. f=frontwheel; r=rearwheel; 4=4wd

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cyl, y = hwy))
```

###Section 3.2.4 5) Scatterplot class vs drv is not useful because both variables are categorical.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = class))
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color='blue'))
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), colour="blue")
```

###Section 3.3.1 1) color is not blue, but red. The code is wrong because the color attribute was written within the aes(X= , Y= ) aesthetic. Color is attribute of the geom_point() so should written within that. 

###Section 3.3.1 2) Categorical variables are manufacturers; model; cyl, year; trans; drv; fl; class. Continuous variables are disp; cty; hwy. In the ?mpg, it explains the variable labels.


```{r}
library(ggplot2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = cty, color = trans)) 
```


```{r}
library(ggplot2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = cty, size = trans))  
```
```{r}
library(ggplot2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = cty, shape= trans))
```
###Sect 3.3.1 3) Color aesthetic works well for categorical and continuous variables. Size will work but it does not look good. Shape aesthetic does not work for categorical variable with more than 6 values.




```{r}
library(ggplot2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = cty, color = trans)) 

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = cty, size = trans))  

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = cty, shape= trans))
```

```{r}
library(ggplot2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = cty, color = class))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = cty, alpha = class))

```
###Sect 3.3.1 4) Mapping same variables with multiple aesthetics allows you to see which is better way to display your data.

```{r}
library(ggplot2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = cty, stroke=0.5))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = cty, stroke=2))

```

###Sect 3.3.1 5) Stroke aesthetic allows you to change the size of your dots in scatterplot.


```{r}
library(ggplot2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=trans, y=cty, color = displ<5))
```
###Sect 3.3.1 6) Nothing happens if you only have aesthetic characteristic without a variable. You need to have variables for X and Y for a plot.

```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))+
  facet_wrap(~class,nrow=2)

ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy, color=class))
```
###Sect 3.5.1 4)  Facet_wrap allows you to display multiple graphs. Having multiple graphs is very useful to see each class of data side-by-side, rather than using the color aesthetic to differentiate the data points on one graph. The disadvantage is that it is harder to see the cluster or trend of all data when they are disaggregated on individual graphs. In larger data sets, I would run scatterplots by class as well as by the entire data set to see the clusters and trends. 



```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))+
  facet_wrap(~class,ncol=10)

ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ, y=hwy))+
  facet_grid(~class)
```

###Sect 3.5.1 5) nrow defines the number of rows for X; ncol defines the number of columns for X.  You can control the order in which the panels appear. You can have the scales vary with each panel. Facet_grid () gives matrix of panels defined by row and column facetting variables already. You use it when you have two discrete variables, and all combinations of the variables already exist in the data.



```{r}
ggplot(data=mpg)+
  geom_line(mapping=aes(x=hwy, y=cty))
  
```
```{r}
ggplot(data=mpg)+
geom_boxplot(mapping=aes(group=class, x=hwy, y=cty))
```
```{r}

ggplot(aes(x=cyl), data=mpg) +
    geom_histogram()
```
```{r}
ggplot(data=mpg)+
  geom_area(mapping=aes(x=cyl, y=cty))
```
###Sec 3.6.1 1) geom_line; geom_boxplot; geom_histogram; geom_area

```{r}
ggplot(data=mpg,mapping=aes(x=displ, y=hwy, color=drv))+
  geom_point()+
  geom_smooth(se=FALSE)
```



```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = trans))
  
    show.legend = FALSE
```
###Sec 3.6.1 3) show.legend= FALSE removes the legend for categorical variables. Not including in the coding defaults to having a legend. Use this code when you don't want a legend.


```{r}
ggplot(data=mpg,mapping=aes(x=displ, y=hwy, color=drv))+
  geom_point()+
  geom_smooth(se=TRUE)
```
###Sect 3.6.1 4) geom_smooth (se= ) allows you add or not add confidence shade around the line

```{r}
ggplot(data=mpg,mapping=aes(x=displ, y=hwy)) + 
  geom_point() +
  geom_smooth()

ggplot()+
  geom_point(data=mpg, mapping =aes(x=displ, y=hwy))+
  geom_smooth(data=mpg,mapping =aes(x=displ, y=hwy))
```
###Sect 3.6.1 5) Plots are not different because the code is the same.The first set of code is more efficient.
```{r}
ggplot(data=mpg) +
  geom_col (mapping=aes (x=class, y=cty))

ggplot(data = mpg) +
  geom_bar(mapping = aes(x =class, y=cty), stat="identity")

```
```{r}
ggplot(data=mpg,mapping=aes(x=displ, y=hwy))+
  geom_point()+
  geom_smooth(se=FALSE)

ggplot(data=mpg,mapping=aes(x=displ, y=hwy, group=drv))+
  geom_point()+
  geom_smooth(se=FALSE)
  show.legend=FALSE
  
  ggplot(data=mpg,mapping=aes(x=displ, y=hwy, color=drv))+
  geom_point()+
  geom_smooth(se=FALSE)

  ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv)) + 
  geom_smooth(se=FALSE)

  
  ggplot(data=mpg,mapping=aes(x=displ, y=hwy, linetype=drv, color=drv))+
  geom_point()+
  geom_smooth(se=FALSE)
 
  
```
###Sect 3.6.1 6) Extra credit plots as shown above.

###Sect 3.7.1 2) Geom_col() does not differ from geom_bar() in terms of the output here as coded. Geom_bar() makes the height of the bar proportional to the number of cases in each group. Use geom_col() if you want the heights of the bars to represent values in the data. Geom_bar() uses stat_count by default: it counts the number of cases at each x position. Geom_col() uses stat_identity: it leaves the data as is.

