---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Quynh Tran"
date: "February 27, 2018"
output: 
  html_document: 
    toc: yes
    toc_depth: 2
---
# https://github.com/mightyquynh/compscix-415-2-assignments

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#The tidyverse packages (3 points)


```{r}
library(tidyverse)
sessionInfo()

```

##1) Can you name which package is associated with each task below? 
###Plotting: ggplot2
###Data munging/wrangling - dplyr
###Reshaping (speading and gathering) data - dplyr
###Importing/exporting data - readr

## 2) Now can you name two functions that you’ve used from each package that you listed above for these tasks?
###Plotting - geom_boxplot() geom_line()
###Data munging/wrangling - filter() select()
###Reshaping data - spread() gather()
###Importing/exporting data (note that readRDS and saveRDS are base R functions) - read_delim() write_csvc()


#R Basics (1.5 points)
## 1) Fix this code with the fewest number of changes possible so it works. My_data.name___is.too0ooLong! <- c( 1 , 2   , 3 ) Just removing ! will make it work, although the data name is very bad:


```{r}


My_data.name___is.too0ooLong <- c( 1 , 2   , 3 )

My_data.name___is.too0ooLong

```

## 2) Fix this code so it works. c has to be small cap, and 'it' was missing a closing '.

```{r}
y_string <- c('has', 'an', 'error', 'in', 'it')

y_string

```

## 3) Look at the code below and comment on what happened to the values in the vector. A vector must have elements of the same type. While 1, 2, 5 were entered as numerics, and '3' and '4' as characters, it made all of them characters. 

```{r}
my_vector <- c(1, 2, '3', '4', 5)
my_vector

typeof(my_vector)
```



#Data import/export 

```{r}

file_path <- '~/compscix-415-2-assignments/rail_trail.txt'
rail_trail <- read_delim(delim = '|', file = file_path)

glimpse(rail_trail)
```


```{r}
saveRDS(rail_trail, file = "~/compscix-415-2-assignments/rail_trail.rds")

rail_trail<- readRDS(file ='~/compscix-415-2-assignments/rail_trail.rds')

glimpse(rail_trail)
```


# Visualization
##1) The Mrs. President pie charts would be better as bar charts. Whole colored pies represent 100% but the yes/no values are separated as separate pies, rather than parts of the pie to represent 100%. Even if the yes/no values were represented within one pie, one problem is that the yes/no sum don't total 100%, so it would be misleading still. It should also be 2 separate graphics, age is a different variable from gender. Even though the gender pie is a different color that the age pies, it's lined up as a group.

 
```{r}

library(forcats)
president <- tribble(
  ~age,     ~Agree, ~value, 
  "under45",  "yes",  79, 
   "under45", "no",   16,
  "b45-64",   "yes", 69, 
  "b45-64",  "no",   22,
  "65over", "yes", 44,
  "65over", "no", 39,
)

president2 <-tribble(
  ~gender, ~Agree, ~value,
  "Men", "yes", 65,
  "Men",  "no", 25,
  "Women", "yes",72,
  "Women", "no", 20,
)
president <- president %>% mutate(age = fct_relevel(age, 'under45', 'b45-64', '65over'))
president


ggplot(data=president)+
  geom_bar(mapping=aes(x=age, y=value, fill=Agree), position="dodge", stat="identity") + 
labs(x="Age group", y="Percentage", title="% respondents who agree a woman will be president in their lifetime")

ggplot(data=president2)+
  geom_bar(mapping=aes(x=gender, y=value, fill=Agree), position="dodge", stat="identity") + 
labs(x="gender", y="Percentage", title="% respondents who agree a woman will be president in their lifetime")

```



counts <- table(mtcars$vs, mtcars$gear)
barplot(counts, main="Car Distribution by Gears and VS",
  xlab="Number of Gears", col=c("darkblue","red"),
 	legend = rownames(counts), beside=TRUE)





## 2) Reproduce diamonds graphics

## 3) Make one change on the code and make graphics more useful. More useful would be to sort by the median of diamond cut.

```{r}
library(tidyverse)

ggplot(data = diamonds, mapping = aes(x = cut, y = carat, fill=color))+
  geom_boxplot(position='identity')+ 
  coord_flip() +
  xlab("CUT OF DIAMOND")+
  ylab("CARAT OF DIAMOND")

ggplot(data = diamonds, mapping = aes(x = reorder(cut, carat, FUN=median), y = carat, fill=color))+
  geom_boxplot(position='identity')+ 
  coord_flip() +
  xlab("CUT OF DIAMOND")+
  ylab("CARAT OF DIAMOND") 
  

ggplot(data = diamonds)+
  geom_bar(aes(x=cut,y=price),stat = "summary",fun.y="mean",alpha=1,fill="blue")+
  xlab("Cut Type")+
  ylab("Average Price")+
  ggtitle("Cut vs Price")


```

# Data munging and wrangling 
## 1) Tidy table2, by spreading, it pulled the "type" of cases and populations into new variable.

```{r}
table2

table2 %>%
spread(key=type, value=count)
```

## 2) diamonds with price per carat

```{r}
library(dplyr)

mutate(diamonds, price_per_carat = price/carat)

diamonds

```

## 3) Number of diamonds price > 1000 and carat < 1.5 by cut.

```{r}
diamonds %>%
filter(diamonds, price > 1000 , carat < 1.5) %>%
  group_by (cut) %>%
  summarise 

count(n())
   

```
# What proportion of diamonds price > 1000 and carat < 1.5 by cut

```{r}
diamonds_new%>%
filter(diamonds, price > 1000 & carat < 1.5) %>%
  group_by (cut) %>%
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = ..count../sum(..count..)))
```

   
diamonds %>% 
  group_by(cut) %>% 
  summarise(prop_cut = mean(price > 1000 , carat < 1.5))
  
  
diam_cut <-filter(diamonds, price > 1000 , carat < 1.5) %>%
   count(n())
  
  
  
  diam_cut <-filter(diamonds, price > 1000 , carat < 1.5) %>%
   count(n())
   
diam_cut
summarise(diam_cut, delay = mean(, na.rm = TRUE))

diam_cut <- select (diamonds, price > 1000 , carat < 1.5) %>%
   group_by(diam_cut,cut)
diam_cut <- filter(data=diamonds, price > 1000, cut )
diamonds %>%
  filter (diamonds_new)
  

group_by(cut)
summarise(proportion = mean(dep_delay, na.rm = TRUE))


library(nycflights13)

flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)
mutate(flights_sml,
  gain = arr_delay - dep_delay,
  speed = distance / air_time * 60
)



# EDA txhousing data

## 1) During what time period is this data from? 2000-2015

```{r}
library(ggplot2)
library (dplyr)
View(txhousing)


```


## 2) How many cities are represented? 46 cities

```{r}
txhousing %>% 
  group_by(city) %>% 
  summarise(
    count = n())

```


## 3) Which city, month and year had the highest number of sales? Houston, July 2015 had 8945 sales.

```{r}

txhousing %>% 
  group_by(city, month, year) %>% 
  summarise(sales = mean(sales, na.rm = TRUE)) %>%
 arrange(desc(sales))
```

## 4) What kind of relationship do you think exists between the number of listings and the number of sales? There is a positive relationship between listings and sales. This relationship holds true for the 15-year period.

```{r}
ggplot(data = txhousing) + 
  geom_point(mapping = aes(x = listings, y = sales)) +
  geom_smooth(mapping = aes(x = listings, y = sales))


ggplot(data = txhousing) + 
  geom_point(mapping = aes(x = listings, y = sales)) +
facet_wrap(~ year, nrow = 3)

ggplot(data = txhousing) + 
  geom_point(mapping = aes(x = listings, y = sales)) +
facet_wrap(~ city)


```

Check your assumption and show your work.
## 5) What proportion of sales is missing for each city?

```{r}
sales_missing <- txhousing %>% 
  filter(is.na(sales))

sales_missing %>% group_by(city) %>% summarize(missing_count = n())

sales_missing
  mutate(cancelled = is.na(dep_delay) | is.na(arr_delay))
```




sales_missing <- txhousing %>% group_by(city) %>% summarize(sales_missing = sum(sales, na.rm = FALSE), prop_missing = mean(sales_missing), na.rm = FALSE)

sales_missing <- txhousing %>% group_by(city) %>% summarize(sales_missing = sum(is.na, na.rm = FALSE), prop_missing = mean(sales_missing), na.rm = FALSE)

```{r}
library(dplyr)

ggplot(data = txhousing) +
  geom_bar(mapping = aes(x = city, y = ..count../sum(..count..)))
 
```
filter(data=txhousing(is.na(city))


```

## 6 Looking at only the cities and months with greater than 500 sales:

```{r}
filter(txhousing, sales > 500)
```

###• Are the distributions of the median sales price (column name median), when grouped by city, different? The same? Show your work.

```{r}
filter(txhousing, sales > 500) %>%
  group_by_()
```

###• Any cities that stand out that you’d want to investigate further?
###• Why might we want to filter out all cities and months with sales less than 500?


flights %>%
  group_by(carrier) %>% 
  summarise(arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
 arrange(desc(arr_delay))

min(data=txhousing,sales)


  group_by(city, month, year) %>% 
  summarise(
    min
    
```
flights %>% 
  group_by(city) %>% 
  summarise(
    count = n(),
    count_dist = n_distinct(arr_delay),
    count_non_miss = sum(!is.na(arr_delay))
  )

txhousing %>% 
  group_by(city, month, year) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  min_rank(desc(data=txhousing, sales))
